\section{Week 3 - 24 Jan 2022 - Continuity}
One of the most important reasons for studying metric spaces. Generalises the
notion we saw for real functions. First we recall the definition for the case of
real functions.

\begin{definition}
  Let $f:D\subset\RR\to\RR$ be a function. We say $f$ is continuous at $a\in D$ if 
  \[\forall\eps>0\exists\delta>0 : |x-a|<\delta\implies |f(x)-f(a)|<\eps.\]
  We say $f$ is a continuous function if it is continuous for all $a\in D$.
  
  Notice that we can rewrite the above definition as $|x-a|<\delta \implies
  d_2(x,a)<\delta$, for the standard $\RR$ metric.
  \label{def:realContinuous}
\end{definition}

\begin{definition}
  Let $(X,d_X)$ and $(Y,d_Y)$ be metric spaces. A funciton $f:X\to Y$ is
  $(d_X,d_Y)$-continuous at $a\in X$ if 
  \[\forall \eps>0 \exists \delta>0 : d_X(x,a)<\delta \implies
  d_Y(f(x),f(a))<\eps.\]
  We say $f$ is $(d_X,d_Y)$-continuous if it is $(d_X,d_Y)$-continuous at all
  $a\in\ X$.
  \label{def:metricContinuous}
\end{definition}

\begin{proposition}
  Let $(X,d_X)$ and $(Y,d_Y)$ be metric spaces. Let $f:X\to Y$ be a function
  continuous at $a\in X$ iff 
  \[\forall \eps>0\exists \delta>0 : f(B_X(a,\delta))\subset B_Y(f(a),\eps).\]
  \label{<+label+>}
\end{proposition}
\begin{proof}
  TODO.
\end{proof}

\begin{proposition}
  Let $(X,d_X)$ be a discrete metric space and $(Y,d_Y)$ is any metric space.
  Then any function $f:X\to Y$ is $(d_X,d_Y)$-continuous.
  \label{prop:discreteFuncsContinuous}
\end{proposition}
\begin{proof}
  Let $a\in X, \eps>0$. If we take $\delta=1$, by the defintion of continuity we
  have 
  \[d_X(x,a)<1 \implies d_Y(f(x),f(a))<\eps\]
  Note that $d_Y(f(x),f(a))=0$ since the discrete metric will be less than $1$
  iff it is $0$, i.e. $x=a$.
\end{proof}

\begin{proposition}
  Let $(X,d_X)$ and $(Y,d_Y)$ be metric spaces. Then any constant function
  $f:X\to Y$ is $(d_X,d_Y)$-continuous.
  \label{prop:constContinuity}
\end{proposition}
\begin{proof}
  TODO.
\end{proof}

\begin{example}
  Let $f:\RR^2\to\RR: (x,y)\mapsto x+y$ is continuous wrt the euclidean metric.
\end{example}
\begin{proof}[Solution]
  TODO. Find bounds for $|x+y-(a+b)|$ -- by triangl inequality -- and find
  $\delta$ from there.
\end{proof}
\begin{exercise}
  Show that the function $m:\RR^2\to\RR:(x,y)\to xy$ is continuous wrt the
  Euclidean metric.
  \label{ex:W3L1LastExercise}
\end{exercise}
\begin{proof}[Solution]
  TODO.
\end{proof}
