\section{Lecture 12 - 18 Oct 2021}
On the 3rd isomorphism theorem. Last lecture about group theory. From next lecture on
we'lll talk about group actions. 
\subsection{Third Isomorphism Theorem}
\begin{lemma}[Sheet 2, Q5]
  Let $\phi:G\to G'$ be a homomorphism and $H'<G'$. We claim that the preimage
  $\phi^{-1}(H')<G$ is a subgroup of $G$. 
  \label{lem:preImgSubgroup}
\end{lemma}
\begin{proof}
  Note that we have $1_{G'}\in H'$ since $H'$ is a subgroup of $G'$, so
  $\phi(1_{G})=1_{G'}$ and it follows tht $1_{G}\in \phi^{-1}(H')$. Next we claim that for
  any $x,y\in\phi^{-1}(H')$ we have $xy^{-1}\in\phi^{-1}(H')$. Note that we have, for any
  $h,h'\in H'$, $h'h^{-1}\in H'$ since $H'$ is a subgroup. Moreover, we have that
  $h=\phi(x),h'=\phi(y)$ for some $x,y,\in\phi^{-1}(H')$, hence it follows that
  $\phi(x)\phi(y)^{-1} = \phi(xy^{-1})\in H'$, or equivalently $xy^{-1}\in\phi^{-1}(H')$,
  as required.
\end{proof}

\begin{lemma}
  Let $\phi:G\to G'$ be a surjective homomorphism and $H'<G'$. We claim 
  \[H'\trianglelefteq G' \iff \phi^{-1}(H')\trianglelefteq G\]
  \label{lem:preImgNormSub}
\end{lemma}
\begin{proof}
  Let $g\in G$ and $h\in \phi^{-1}(H')$. Then define $g'=\phi(g)$ and $h'=\phi(h)\in H$.
  We have $g'h' = \bar{h}'g'$ for some $\bar{h}'\in H$
  \[g'h' = \bar{h}'g' \iff \phi(g)\phi(h)  =\phi(\bar{h})\phi(g) \exists
  \bar{h}\in\phi^{-1}(H') \iff \phi(ghg^{-1}) = g'h'(g')^{-1} \in H'\]
  \[\iff ghg^{-1}\in \phi^{-1}(H')\]
  Hence $\phi^{-1}(H')\trianglelefteq G$, as required.
\end{proof}

\begin{theorem} [Third Isomorphism Theorem]
  Let $G$ be a group, and $N\trianglelefteq G$. Define the quotient map $\pi:G\to G/N$.
  Then
  \begin{enumerate}
    \item For every subgroup $U$ of $G/N$, the preimage $\pi^{-1}(U)$ is a subgroup of
      $G$.
    \item If $U$ is a subgroup of $G/N$, then $U$ is normal in $G/N$ if and only iff
      $\pi^{-1}(U)$ is normal in $G$.
    \item The function $U\mapsto\pi^{-1}(U)$ defines a bijection between the set of
      subgroups of $G/N$ and the set of subgroups of $G$ that contain $N$.
    \item If $K\trianglelefteq G$ containing $N$, then $\pi$ defines an isomorphism
      \[G/K \cong (G/N)/(K/N)\]
      By the isomorphism $gK\mapsto (gN)(K/N)$.
  \end{enumerate}
  \label{thm:tit}
\end{theorem}
\begin{proof}
  We'll prove these part by part. For the first part, we have that
  $N=\pi^{-1}(1N)\subset\pi^{-1}(U)$ for every subset of $U$, since $U$ is a subgroup of
  $G/N$ it must contain the identity element, naming $N$. Note that the pre-image of $N$
  is $N$ itself (recall the absorption rule from 2F). Moreover, we have that for a
  homomorphism $\phi:G\to G/N$, the preimage of any subgrup of $G/N$ is a subgroup of $G$
  by Lemma \ref{lem:preImgSubgroup}.

  For the second part, we have that the quotient map is surjective by definition. The
  proof then follows by Lemma \ref{lem:preImgNormSub}.

  For the third part, we find an inverse of the function $U\to \pi^{-1}(U)$ to show that
  it's indeed a bijection. Let $\SH$ be the set of subgroups of $G$ that contain $N$, i.e.
  $\SH=\{H<G | H\cap N = N\}$, and let $\SH'$ be the st of subgroups of $G/N$, i.e.
  $\SH'=\{gN | g\in G\}$. Then we find $\SH\to\SH'$ by $H\mapsto \pi (H)=H/N$. Hence the
  defined function is a bijection.

  For the last part, we check that the isomorphism is well-defined, and that it's
  bijective. To check it's well defined, suppose $gK=g'K$ for some $g,g'\in G$. Then
  $g^{-1}g'\in K$, so $g^{-1}g' N \in K/N$ hence $gN(K/N) = g'N(K/N)\in (G/N)/(K/N)$.
  Surjectivity is given since every coset representative of $G/N$ is hit by some
  representative of $G/K$. Injectivity proof is the same as the well-definedness proof but
  read backwards (notice how the proof goes both ways!)

\end{proof}

The above theorem says that if $G$ is a group, $K$ is a normal subgroup, and $\phi$ is a
homomorphism from $G$ whose kernel ($N$ above, since every normal subgroup is a kernel of
a homomorphism) is contained in $K$, then $G/K \cong \phi(G)/\phi(K)$. Note that
$\phi(G)=\Img \phi:G\to G/N$ and $\phi(K)=\Img \phi_{K}:K\to K/N$. This, in order to
understand $G/K$ we may apply any homomorphism $\phi$ whose kernel ($N$) is contained in
$K$ and instead understand that quotient on the other side of the homomorphism, which is
often simpler.

\begin{example}
  Let $G=\GL_{2}(\RR)$ and let $K=\GL_2^+(\RR)=\{X\in\GL_2(\RR) : \det X>0\}$. Then it's
  easy to see that $K$ is normal in $G$. What is $G/K$? THe gorup $K$ contains the normal
  subgroup $N=\SL_2(\RR)=\ker\det$. We have surjective group homomorphisms
  $G\to\RR^{\times}$ and $\GL_2^+(\RR)\to\RR_{>0}$, with kernel $\SL_2(\RR)$ and by the
  FIT we have $G/N\cong \RR^{\times}$ and $K/N\cong \RR_{>0}$. By the TIT we have
  $G/K\cong \det\GL_2(\RR)/\det\GL_2^+(\RR)\cong \RR^{\times}/\RR_{>0}\cong \{\pm 1\}$.
\end{example}

Get used to thinking about normal subgroups as kernels of homomorphisms, and quotients
$G/N$ as images $\phi(G)$ of homomorphisms $\phi$ whose kernels are that normal subgroup
$N=\ker\phi$.
